---
title: "Cross-Validation and Bayes Homework"
author: "j4yr0u93"
date: "10/28/2020"
output: html_document
---

```{r load, include=FALSE}
library(tidyverse)
library(modelr)
library(rsample)
library(AICcmodavg)
library(boot)
library(palmerpenguins)
poggers <- read_csv("../data/progesterone.csv")
```

# 1
```{r polyamorous_plot, echo=FALSE}
#ggplot with standard lm line and poly line
poggers %>% ggplot(mapping = aes(x = ventilation, y = progesterone)) + geom_point() + stat_smooth(method = "lm", formula = y ~ poly(x,3), color = 'purple', alpha = 0.2, fill = 'blue') + stat_smooth(method = "lm", formula = y ~ x, color = 'orange', alpha = 0.2, fill = 'yellow')

#standard lm and poly scale x val lm
prog_lm <- lm(progesterone ~ ventilation, data = poggers)
prog5_lm <- lm(progesterone ~ poly(ventilation, 5), data = poggers)

#rsquare values for each model
rsquare(prog_lm, poggers)
rsquare(prog5_lm, poggers)
```

The 5 poly scale seems to have a higher r-squared value, indicating more of a relationship between the independent variable and dependent variable. \

# 2a
```{r 5_finger_fold_run, echo=FALSE}
#vfold x crossing for all 5 possibilities
pog_folds <- vfold_cv(poggers, 5)
polys <- c(1:5)

pog_fold <- crossing(pog_folds$id, polys)

names(pog_fold)[names(pog_fold) == "pog_folds$id"] = "id"

pog_fold <- pog_fold %>% full_join(pog_folds)

summary(pog_fold)
```

# 2b
```{r polyglot_cats, echo=FALSE}
#map2 fm(fit models) with poly of different possibilities 1:5
pog_cat <- pog_fold %>% mutate(fm = map2(splits, polys,
                    ~lm(ventilation ~ poly(progesterone, .y), 
                        data = analysis(.x))))
summary(pog_cat)
```

# 2c
```{r ramsey, echo=FALSE}
#rmse with pog cat and display summary
pog_rmse <- pog_cat %>% mutate(rmse = map2_dbl(splits, fm, ~rmse(model = .y, data = assessment(.x))))

summary(pog_rmse)
```

# 2d
```{r imps, echo=FALSE}
#plotting mean rmse for polys to visualize best poly for model
pog_imps <- pog_rmse %>% group_by(polys) %>% summarize(mean_rmse = mean(rmse)) %>% ungroup() 

pog_imps %>% ggplot(mapping = aes(x = polys, y = mean_rmse)) + geom_point() + geom_line() + theme_minimal()
```
A poly of 4 gives us our peak rmse so it is likely the optimal model.

# 3
```{r AOC, echo=FALSE}
#AIC for poly pick to compare to mean rmse
pog_kibble <- data.frame(polys = 1:5) %>% mutate(model = map(polys, ~lm(ventilation ~ poly(progesterone, .x), data = poggers)))

aictab(pog_kibble$model, pog_kibble$polys)
```
AICtab says that 5 poly is a better model than 4, which I agree with since my initial plot was a 5 model and to my naked eye the fit does look better.

# 4
```{r cowboy_boot, echo=FALSE}
#booting the poggers to get a cv for each poly
pog_cv <- data.frame(polys = 1:5) %>% mutate(gm = map(polys, ~glm(ventilation ~ poly(progesterone, .x), data = poggers))) %>%
  mutate(cv = map2_dbl(polys, gm, ~cv.glm(poggers, .y)$delta[1])) %>% select(polys, cv)

pog_cv
```

# 5a

```{r pingu, echo=FALSE}
#load peng and plot basic for viz
pingu <- penguins %>% filter(species == 'Gentoo', sex == 'female', !is.na(flipper_length_mm))

pingu %>% ggplot(mapping = aes(x = flipper_length_mm)) +
  geom_density() +
  theme_minimal()
```

# 5b
```{r pingu_grid, echo=FALSE}
#make grid for bayes func ltr
grid_init <- crossing(mean = seq(205, 220, length.out = 100),
                      sd = seq(3, 5, length.out = 100))
#show grid exists properly
str(grid_init)
```

# 5c
```{r happy_meal_combo, echo=FALSE}
#get my numerator
wombo_combo <- function(mean, sd){
  sum(dnorm(pingu$flipper_length_mm, mean, sd, log = TRUE)) +
    dnorm(mean, 210, 50, log = TRUE) +
    dunif(sd, 1, 10, log = TRUE)
}
```

# 5d
```{r pingu_ping, ECHO=FALSE}
#execute function on the data to get log posterior
post <- grid_init %>%
  rowwise(mean,sd) %>%
  mutate(num = wombo_combo(mean,sd)) %>%
  ungroup() %>%
  mutate(log_post = num - log(sum(exp(num))),
         post = exp(log_post))

post %>% filter(log_post == max(log_post))
```

I mean it is a bayes approach rather than frequentist so the difference in the modal estimate is a question of likelihood rather than significance. It is unlikely to pick this particular penguin out of all of our penguins even though visually it is on the peak of our 'normal' curve. Our skew in our data is positive and kurtosis is potentially less than 1. Normal but not perfectly uniform data.