---
title: "Functions and Tidy Data"
author: "j4yr0u93"
date: "10/5/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
sample_vec <- c('orange', 'apple', 'pear', 'banana', 'kiwi', 'starfruit', 'plum')
sample_vec2 <- rnorm(10, mean = runif(1, min = 1, max = 100), sd = runif(1, min = 3, max = 5))
```
# 1
```{r boot_funk, echo=FALSE}
boot <- function(a_vec){
  #get a sample length equal to vector length
  samplen <- round(length(a_vec))
  #take the sample from the vector, with replacement
  a_sample <- sample(a_vec, samplen, replace = TRUE)
  #return the sample
  return(a_sample)
}

boot(sample_vec)
```
# 2
```{r stat_funk, echo=FALSE}
stat_boot <- function(a_vec, R_num = 1000, a_funk = 'mean'){
  #convert a_funk string into the actual funk
  funk <- get(a_funk)
  #create first 'row' of stat_df to self rbind in loop
  stat_df <- data.frame(stat = funk(boot(a_vec)))
  #loop for R_num iterations
  while(nrow(stat_df) < R_num){
    stat_df <- rbind(stat_df, data.frame(stat = funk(boot(a_vec))))
  }
  #rename column to funk used
  colnames(stat_df) <- a_funk
  #return the df of funks from samples
  return(stat_df)
}

stat_boot(sample_vec2, 5)
```
# 3
```{r more_stats, echo=FALSE}
stat_boot_more <- function(a_vec, R_num = 1000, a_funk = 'mean'){
  #funk get
  funk <- get(a_funk)
  #get the basic df to work with
  stat_df <- stat_boot(a_vec, R_num, a_funk)
    #add my mutations to it
  stats_df <- data.frame(og_stat = funk(a_vec),
                         boot_stat_mean = mean(stat_df[,1]),
                         upper_CI = quantile(stat_df[,1], probs = 0.975),
                         lower_CI = quantile(stat_df[,1], probs = 0.025),
                         bias = funk(a_vec) - mean(stat_df[,1]),
                         row.names = NULL)
  return(stats_df)
}

stat_boot_more(sample_vec2, 50, 'mean')
```

```{r load_poll, include=FALSE}
#get data
prez_pol <- read_csv('../data/president_polls.csv')
#load lub for mdy()
library(lubridate)
```
# 4a
```{r pol_peek, echo=FALSE}
#peek pol
head(prez_pol)
```
\
data appears to be hybrid \

# 4b
```{r date_filter, echo=FALSE}
#lub filter for prez
current_prez_pol <- prez_pol %>% filter(answer %in% c('Trump','Biden'), mdy(start_date) > mdy('9/29/20'))
#peek pol
head(current_prez_pol)
```

# 4c
```{r avg_polls, echo=FALSE}
#get booted national avgs
national_avg <- current_prez_pol %>% filter(is.na(state)) %>% group_by(answer) %>% mutate(pol_avg = stat_boot(pct, 1)$mean) %>% select(answer, pol_avg) %>% distinct()
national_avg
```

# 4d
```{r state_dif_boot, echo=FALSE}
#booted diffs between pols by state w/ national
pol_diffs <- current_prez_pol %>%
             mutate(ID = paste(question_id, poll_id, state)) %>%
             select(ID, state, answer, pct) %>%
             pivot_wider(names_from = answer,
              values_from = pct) %>%
             mutate(diff = Biden - Trump) %>%
             filter(!is.na(diff)) %>%
             group_by(state) %>%
             summarize(stat_boot(diff,1)$mean)

pol_diffs[is.na(pol_diffs)] <- 'National'
pol_diffs
```

# 5a
```{r finding_sd_mean, echo=FALSE}
#find mean & sd and then rnorm to make vals
biden_nat_sd <- current_prez_pol %>% filter(answer == 'Biden', is.na(state)) %>% mutate(nat_sd = sd(pct)) %>% select(nat_sd) %>% distinct() %>% as.double()
biden_nat_mean <- current_prez_pol %>% filter(answer == 'Biden', is.na(state)) %>% mutate(nat_mean = mean(pct)) %>% select(nat_mean) %>% distinct() %>% as.double()
biden_sim <- replicate(1000, rnorm(50, mean = biden_nat_mean, sd = biden_nat_sd))
```
\
We get an extra large and nasty matrix thing of 50 by 1000 vals, I don't like it. \

# 5b
```{r tibblify, echo=FALSE}
#pivot long to get nice sims
biden_sim_clean <- biden_sim %>%
  as_tibble(.name_repair) %>%
  pivot_longer(cols = everything(),
               names_to = "sim",
               values_to = "pct")
#peek sims
head(biden_sim_clean)
```
\
This looks way better pivoting it just peeking it. Actually seems readable, not that I would read 50k entries. \
\

# 5c
```{r bidens_big_boot, echo=FALSE}
#simple groupby and summarize with funk 3 and then select/distinct because I dislike extra cols
bidens_boot <- biden_sim_clean %>%
  group_by(sim) %>%
  summarize(stat_boot_more(pct, 50)) %>%
  select(sim, og_stat, boot_stat_mean, upper_CI, lower_CI) %>%
  distinct()

#plotting boot and sample means with lines showing means of both and red lines with the actual population mean
ggplot(bidens_boot, mapping = aes(x = og_stat, y = boot_stat_mean, ymin = lower_CI, ymax = upper_CI)) +
  geom_pointrange(alpha = 0.2) +
  geom_vline(xintercept = mean(bidens_boot$og_stat)) +
  geom_hline(yintercept = mean(bidens_boot$boot_stat_mean)) +
  geom_vline(xintercept = biden_nat_mean, color = 'red', alpha = 0.5) +
  geom_hline(yintercept = biden_nat_mean, color = 'red', alpha = 0.5)

boot_mean_greater_count <- length(bidens_boot$boot_stat_mean[bidens_boot$boot_stat_mean > biden_nat_mean])
sample_mean_greater_count <- length(bidens_boot$og_stat[bidens_boot$og_stat > biden_nat_mean])

print(paste('A is', boot_mean_greater_count, 'and B is', sample_mean_greater_count))
```
\
The bootstrapped mean was greater than original A times. The sample mean was greater than the original B times. \

# 5d
\
The black intercepts above are means of each axis and the red lines are both the actual means from the pol (visually see difference between boot mean and og mean, as well as sample mean and og mean). Something weird might be going on. CIs formed the lines on each point which just shows the proximal range the simulation occupied for a given point. They also indicate if a simulation contained the true mean when seeing if the red line falls on them. \

# cool current biden map
```{r eyes_bleeding_code, include=FALSE}
#lib loading
library(ggmap)
library(sf)
library(USAboundaries)

#US map data loading
us_states <- us_states() %>%
  filter(name != "Alaska") %>%
  filter(name != "Hawaii")

#getting state names for nested loop
states <- unique(current_prez_pol$state) %>% na.omit()

#wrapper function because I could, literally could have just used groupby
get_state_avgs <- function(){
  state_avg <- data.frame()

  for (i in states){
    for (j in c('Biden', 'Trump')){
      a_state <- current_prez_pol %>% filter(answer == j, state == i) %>%
                                      mutate(pol_avg = mean(pct),
                                      state_nat_diff = pol_avg - filter(national_avg, answer == j)$pol_avg) %>%
                                      select(state, answer, pol_avg, state_nat_diff) %>% distinct()
      state_avg <- rbind(a_state, state_avg)
    }
  }

  return(state_avg)
}

#execute wrapper
state_avg <- get_state_avgs()

#extract biden info
biden_state_nat_diff <- state_avg %>% filter(answer == 'Biden') %>% rename(name = state)

#join map data and biden data
biden_diff_map <- full_join(biden_state_nat_diff, us_states) %>% st_as_sf(crs = 4326)
```

```{r plotty, echo=FALSE}
ggplot(biden_diff_map) +
  geom_sf(us_states, mapping = aes(fill = FALSE)) +
  geom_sf(biden_diff_map, mapping = aes(fill = state_nat_diff)) +
  scale_fill_gradient(low = "blue", high = "orange", trans = 'reverse')
```
\
Just a map visualizing state differences in polling from the national for Biden approval/support. Orange states have less than national average and blue or purple state have higher than national average. Note that the range of avg state pols compared to national is only about 7 points in both directions. \

# git extra credit
pushed my repo and submitting the whole repo as hw \
[HW_4 github link](https://github.com/j4yr0u93/HW_4)