---
title: "GLM Practice"
author: "j4yr0u93"
date: "11/25/2020"
output: html_document
---

```{r setup, include=FALSE}
#load libs and data
library(tidyverse)
library(readr)
library(ggfortify)
library(car)
library(emmeans)
library(janitor)
library(brms)
library(bayesplot)
library(tidybayes)
library(ggdist)
cone <- read_csv("../data/PineCones.csv")
#make cone model
cone_lm <- lm(conemass ~ habitat, data = cone)
```

# 1.1
```{r 1.1, echo=FALSE}
#boxplot cause lazy & easy
ggplot(cone, mapping = aes(x = habitat, y = conemass)) +
  geom_boxplot() +
  theme_minimal()
```

# 1.2
```{r 1.2, echo=FALSE}
#test assumptions
par(mfrow=c(1, 3))
plot(cone_lm, which = c(1, 2, 4))

#resid plot
residualPlots(cone_lm)

#lev test
leveneTest(cone_lm)
```
 \
Residuals vs fitted, qq plots, and cook's distance all look normal within acceptable range and linear.


# 1.3
```{r 1.3, echo=FALSE}
#r square check
summary(cone_lm)$r.squared
```
 \
~89% of variance is explained by the independent variable.

# 1.4
```{r 1.4, echo=FALSE}
#check emmeans to see diff and outlier
emmeans(cone_lm, ~habitat) %>%
  plot() +
  theme_minimal()
```
 \
Island Absent in the outlier from the means, non-default correction not required. \

# 2.1
```{r 2.1, echo=FALSE}
#load trans data
transplant <- read_csv('../data/fouling_transplant_data.csv') %>% clean_names()

#facet for combos and plot to see basic trend
transplant %>% 
  ggplot(aes(x = initial_cover, y = final_cover)) +
  geom_boxplot() +
  facet_wrap(position_on_block~caged)
```

# 2.2
```{r 2.2, echo=FALSE}
#likelihood model
trans_glm <- glm(change_in_cover ~ caged * position_on_block,
                 data = transplant,
                 family = gaussian())

#plot relevant tests
par(mfrow=c(1,3))
plot(trans_glm, which = c(1, 2, 4))

#shappy test
shapiro.test(residuals(trans_glm))
#lev
leveneTest(trans_glm)
```
 \
Residuals vs fitted, qq plots, and cook's distance all look normalish within acceptable range and linear. Separation of points on resid explained by caged * position_on_block. Shappy test and Lev test show that the model is not meeting assumptions.

# 2.3
```{r 2.3, echo=FALSE}
#covariance w/ shap
trans_cov <- glm(final_cover ~ caged * position_on_block + initial_cover,
                 data = transplant,
                 family = gaussian())

shapiro.test(residuals(trans_cov))

#change model w/ shap
trans_delta <- glm(change_in_cover/initial_cover ~ caged * position_on_block,
                   data = transplant,
                   family = gaussian())

shapiro.test(residuals(trans_delta))

#logit model w/ shap
trans_logit <- glm(logit(initial_cover) - logit(final_cover) ~ caged * position_on_block,
                   data = transplant,
                   family = gaussian())

shapiro.test(residuals(trans_logit))

#lev on relevant model
leveneTest(trans_delta)

#plot relevant tests
par(mfrow=c(1,3))
plot(trans_delta, which = c(1, 2, 4))
```
 \
trans_delta or option 2 has a larger p val and slightly wack qq/resid plots but the best cooks vals. Also has the smallest AIC val which correlates to  better fit. \

# 2.4
```{r 2.4, echo=FALSE}
#anova relevant model
anova(trans_delta)

#emeans checks of possible variances in model
emmeans(trans_delta, ~caged) %>%
  contrast(method = "tukey")

emmeans(trans_delta, ~position_on_block) %>%
  contrast(method = "tukey")

#plot 2 options
par(mfrow=c(2,1))

emmeans(trans_delta, ~caged) %>%
  as_tibble %>%
  ggplot(aes(x = caged, y = emmean, 
             ymin = emmean-SE, ymax = emmean+SE)) +
           geom_pointrange() +
  theme_minimal()




emmeans(trans_delta, ~position_on_block) %>%
  as_tibble %>%
  ggplot(aes(x = position_on_block, y = emmean, 
             ymin = emmean-SE, ymax = emmean+SE)) +
           geom_pointrange() +
  theme_minimal()

```
 \
Difference in position has same diff in mean as caged. Impacts on the system are equal and connected for open vs caged. Both variables appear to be significant with alpha 0.08 on the systems and signif sub that alpha. \

# 3.1
```{r 3.1, echo=FALSE}
#load rat
rat <- read_csv('../data/MoleRat.csv') %>% clean_names()

#prelim plot wrap
ggplot(rat,
       aes(x = lnmass,
           y = lnenergy)) +
  geom_point() +
  stat_smooth() +
  facet_wrap(.~caste) +
  theme_minimal()

#rat bay
rat_brm <- brm(lnenergy ~ lnmass + caste,
               data = rat)

#check rat bay
pp_check(rat_brm)
```
 \
BRM pp_check looks like normal chains with somewhat of a normal curve. Data should meet assumptions.\

# 3.2
```{r 3.2, echo=FALSE}
#interaction rat
rat_brm2 <- brm(lnenergy ~ lnmass * caste,
               data = rat)

#compare rats
loo_compare(loo(rat_brm), loo(rat_brm2))
```
 \
First brm is better option for modeling, interaction model is messy. lnmass + caste > lnmass * caste. \

# 3.3
```{r 3.3, echo=FALSE}
#emmeans looking at lazy vs. worker
emmeans(rat_brm, ~caste) %>%
  as_tibble() %>%
  ggplot(aes(x = caste, y = emmean, 
             ymin = lower.HPD, ymax = upper.HPD)) +
           geom_pointrange() +
  theme_minimal()
```
 \
There is a clear difference in the means of lazy and worker. I would plot my model before discussing my conclusions further. I assume a more visually appealing plot will make it more clear. \

# 3.4
```{r 3.4, echo=FALSE}
#very pretty plot
rat %>%
  group_by(caste) %>%
  modelr::data_grid(lnmass = modelr::seq_range(lnmass, 100)) %>%
  ungroup() %>%
  add_fitted_draws(rat_brm) %>%
  rename(lnenergy = .value) %>%
  ggplot(aes(x = lnmass, y = lnenergy, color = caste)) +
  stat_lineribbon() +
  scale_fill_viridis_d() +
  geom_point(data = rat) +
  theme_minimal()
```